Incorporating periodic variability in hidden Markov models for animal movement

### 2016-07-15 paper3.Rnw
update figure 3 ...

<<adj_BIC_comparisons, warning=FALSE, echo=FALSE, message=FALSE,fig.cap="Overall adjusted BIC comparisons among models, cat \\#1. \\textbf{todo: use dashed lines for FMMs (leadup to next figure)?}",fig.width=8>>=
#L <- load("cat1sumdat.RData")
sumdat <- transform(readRDS("summary_stats/cat.1.df.RDS")$HMMsumdf,
                    mod = ifelse(grepl("HMM",model),"no","yes")
)
linetypev <- c(rep(2,2),rep(1,5))
(adj_BIC_plot <- ggplot(sumdat, aes(x=nstates
                                    , y=deltaBIC
                                    , colour=model
                                    , linetype=model))
+ facet_wrap(~ type,ncol=4)
+ scale_size_continuous( name="# of parameters")
+ scale_x_continuous(expand=c(0,0.5),breaks=2:7)
+ geom_point(aes(size=parameters))
+ scale_linetype_manual(values=linetypev)
+ scale_colour_brewer(palette="Dark2")
+ labs(y=expression(paste(Delta, "BIC"))
       , x="Number of States") 
+ geom_line() 
+ zmargin)

@@

# sumdat <- transform(readRDS("summary_stats/cat.1.df.RDS")$HMMsumdf,
#                     mod = ifelse(type=="HMM + TH","HMM + TH","FMMs and HMM")
# )
# linetypev <- c(rep(2,2),rep(1,5))
# (adj_BIC_plot <- ggplot(sumdat, aes(x=parameters
#                                     , y=deltaBIC
#                                     , colour=model
#                                     , linetype=model
#                                     , group=model))
#   + facet_wrap(~mod,ncol = 2,scales = "free_x")
#   + scale_size_continuous(name="# of states")
#   + scale_x_continuous(breaks=seq(0,300,by=10))
#   + geom_point(aes(shape=factor(nstates),size=1.5,alpha=nstates))
#   + scale_linetype_manual(values=linetypev)
#   + scale_colour_brewer(palette="Dark2")
#   + labs(y=expression(paste(Delta, "BIC"))
#          , x="Number of States") 
#   + geom_line()
#   + zmargin)
@

@@

combdf <- (rbind(avgdf,obsdf)
           %>% transmute(time=time,model=variable,steplen=value)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)
linetypev <- c(rep(c(2,1),c(2,6)),1)
colvec <- c(c(brewer.pal(7,"Dark2"),"black"),"black")
lwdvec <- c(rep(c(1,4),c(7,1)),4)
alphavec <- c(rep(c(1,0.2),c(7,1)),1)
(ggplot(combdf,aes(x=time,y=steplen,colour=model,lwd=model,alpha=model,linetype=model)) 
  + geom_line()
  + xlab("Time of Day") 
  + ylab(expression('Mean step length'~(log[10]*m)))
  + scale_color_manual(values=colvec)
  + scale_alpha_manual(values=alphavec)
  + scale_size_manual(values=lwdvec)
  + scale_linetype_manual(values=linetypev)
) + geom_point(data=vitdf,aes(y=viterbi),lwd=2.5,alpha=1,colour="black",shape=118)
@

@@

acfdf <- (readRDS("summary_stats/cat.1.df.RDS")$acfdf 
          %>% mutate(model = .id)
)
(ggplot(acfdf,aes(x=lag,y=ACF,colour=model,lwd=model,alpha=model,linetype=model)) 
  + geom_line()
  + xlab("Lag") 
  + ylab("Autocorrelation")
  + scale_color_manual(values=colvec)
  + scale_alpha_manual(values=alphavec)
  + scale_size_manual(values=lwdvec)
  + scale_linetype_manual(values=linetypev)
) + geom_point(data=vitacf,lwd=2.5,alpha=1,colour="black",shape=118)
@

>>>>

<<adj_BIC_comparisons, warning=FALSE, echo=FALSE, message=FALSE,fig.cap="Overall adjusted BIC comparisons among models, cat \\#1. \\textbf{todo: use dashed lines for FMMs (leadup to next figure)?}",fig.width=8>>=
#L <- load("cat1sumdat.RData")
sumdat <- transform(readRDS("summary_stats/cat.1.df.RDS")$HMMsumdf,
                    mod = ifelse(grepl("HMM",model),"no","yes")
)
linetypev <- c(rep(2,2),rep(1,5))
## adj_BIC_plot <- ggplot(sumdat, aes(x=nstates
##                                     , y=deltaBIC
##                                     , colour=model
##                                     , linetype=model))
## + facet_wrap(~ type,ncol=4)
## + scale_size_continuous( name="# of parameters")
## + scale_x_continuous(expand=c(0,0.5)) ## ,breaks=2:7)
## + geom_point(aes(size=parameters))
## + scale_linetype_manual(values=linetypev)
## + scale_colour_brewer(palette="Dark2")
## + scale_shape_discrete(name="# of states")
## + scale_alpha_discrete(name="# of states")    
## + labs(y=expression(paste(Delta, "BIC"))
##        , x="Number of States") 
## + geom_line() 
## + zmargin

@@

sumdat <- transform(readRDS("summary_stats/cat.1.df.RDS")$HMMsumdf,
                     mod = ifelse(type=="HMM + TH","HMM + TH","FMMs and HMM")
)
linetypev <- c(rep(2,2),rep(1,5))
(adj_BIC_plot <- ggplot(sumdat, aes(x=parameters
                                  , y=deltaBIC
                                  , colour=model
                                  , linetype=model
                                  , group=model))
   + facet_wrap(~mod,ncol = 2,scales = "free_x")
   + scale_size_continuous(name="# of states")
   + scale_x_continuous(breaks=seq(0,300,by=30))
   + geom_point(aes(shape=factor(nstates),alpha=nstates))
   + scale_linetype_manual(values=linetypev)
   + scale_colour_brewer(palette="Dark2")
   + scale_shape_discrete(name="Number of states")
   + scale_alpha_continuous(name="Number of states")
   + labs(y=expression(paste(Delta, "BIC"))
          , x="Number of parameters") 
   + geom_line()
   + zmargin)
@

@@

combdf <- (rbind(avgdf,obsdf)
           %>% transmute(time=time,model=variable,steplen=value)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)
linetypev <- c(rep(c(2,1),c(2,6)),1)
colvec <- c(c(brewer.pal(7,"Dark2"),"black"),"black")
lwdvec <- c(rep(c(1,4),c(7,1)),4)
alphavec <- c(rep(c(1,0.2),c(7,1)),1)
(ggplot(combdf,aes(x=time,y=steplen,colour=model,lwd=model,alpha=model,linetype=model)) 
  + geom_line()
  + xlab("Time of Day") 
  + ylab(expression('Mean step length'~(log[10]*m)))
  + scale_colour_manual(values=colvec)
  + scale_alpha_manual(values=alphavec)
  + scale_size_manual(values=lwdvec)
  + scale_linetype_manual(values=linetypev)
) + geom_point(data=vitdf,aes(y=viterbi),
               lwd=2.5,alpha=1,colour="black",shape=118,show.legend=FALSE)
## adding geom_point() with vitdf screws up the legend ...
@

@@

acfdf <- (readRDS("summary_stats/cat.1.df.RDS")$acfdf 
          %>% mutate(model = .id)
)
(ggplot(acfdf,aes(x=lag,y=ACF,colour=model,
                  lwd=model,alpha=model,linetype=model)) 
  + geom_line()
  + xlab("Lag") 
  + ylab("Autocorrelation")
  + scale_colour_manual(values=colvec)
  + scale_alpha_manual(values=alphavec)
  + scale_size_manual(values=lwdvec)
  + scale_linetype_manual(values=linetypev)
) + geom_point(data=vitacf,lwd=2.5,alpha=1,colour="black",shape=118,
             )
## vitacf messes up legend ...
@
